A Process for the Emulation of Comparative Oncology Trials with Real-world Evidence (ENCORE)

### 2024-12-14 manuscript/manuscript.qmd
renamed functioin
modified figure 3

@@

source(here("functions", "simulate_flaura.R"))

data <- simulate_flaura(
  n_total = 2000, 
  seed = 41, 
  include_id = FALSE, 
  imposeNA = TRUE,
  propNA = .25
  ) |> 
  mutate(treat = ifelse(treat == 1, "Drug A", "Drug B"))

# impute
mids_data <- mice(data, m = 10, printFlag = F)

@@

# weighting  
wimids_data <- weightthem(
  datasets = mids_data, 
  formula = ps_formula, 
  method = "glm",
  estimand = "ATO"
  )

@@

# covariate balance
covariate_balance <- love.plot(
  x = wimids_data,
  abs = TRUE,
  var.names = named_covariates,
  thresholds = 0.1, 
  drop.distance = TRUE,
  var.order = "unadjusted",
  colors = c("orange", "blue"), 
  stars = "std",
  shapes = 17, 
  size = 4, 
  grid = TRUE,
  position = "top"
  )

# distributional balance
score_balance <- bal.plot(
  x = wimids_data,
  var.name = "prop.score",
  which = "both",
  which.imp = .none,
  colors = c("orange", "blue")
  ) +
  labs(
    x = "Prognostic score",
    title = "Distributional balance for prognostic score"
    ) +
  theme(legend.position = "top")

>>>>

To ensure a transparent, reproducible and consistent way of deriving analytic cohorts and performing statistical analyses across trial emulations, we developed and internal R package `encore.io`with parameterized functions

@@

source(here("functions", "simulate_data.R"))

data <- simulate_data(
  n_total = 2000, 
  seed = 41, 
  include_id = FALSE, 
  imposeNA = TRUE,
  propNA = .25
  ) |> 
  mutate(treat = ifelse(treat == 1, "Drug A", "Drug B"))

# impute
suppressMessages(mids_data <- mice(data, m = 10, printFlag = F))

@@

# weighting  
suppressMessages(wimids_data <- weightthem(
  datasets = mids_data, 
  formula = ps_formula, 
  method = "glm",
  estimand = "ATO"
  ))

@@

# covariate balance
covariate_balance <- love.plot(
  x = wimids_data,
  abs = TRUE,
  thresholds = 0.1, 
  drop.distance = TRUE,
  var.order = "unadjusted",
  colors = c("orange", "blue"), 
  stars = "std",
  shapes = 17, 
  size = 4, 
  grid = TRUE,
  position = "top"
  )

# distributional balance
# Simulate data
set.seed(42)  # For reproducibility
n <- 2000  # Number of samples per group
drugA_pre <- rnorm(n, mean = 0, sd = 2.5 / 2)  # Spread = 2.5
drugB_pre <- rnorm(n, mean = -1, sd = 2.5 / 2)  # Spread = 2.5
drugA_post <- rnorm(n, mean = 0, sd = 2.5 / 2)  # Spread = 2.5
drugB_post <- rnorm(n, mean = 0, sd = 2.5 / 2)  # Spread = 2.5

# Combine into a data frame
data <- data.frame(
  Value = c(drugA_pre, drugA_post, drugB_pre, drugB_post),
  Group = rep(c("Drug A", "Drug A", "Drug B", "Drug B"), each = n),
  PrePost = rep(c("Unadjusted sample", "Adjusted sample", "Unadjusted sample", "Adjusted sample"), each = n)
  ) |> 
  mutate(PrePost = factor(PrePost, levels = c("Unadjusted sample", "Adjusted sample")))

# Create the plot
score_balance <- ggplot(data, aes(x = Value, fill = Group)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = c("orange", "blue")) +
  labs(
    title = "Distributional balance for prognostic score",
    x = "Value",
    y = "Density",
    fill = "Group"
  ) +
  theme_minimal() +
  theme(legend.position = "top") +
  facet_wrap(~PrePost)
